---
title: "Importing and Wrangling Data"
author: Genee' Glascoe
date: "April 16, 2020" 
output: html_notebook
---

**This notebook is to eplore about importing and wrangling data utilzing tidyverse**

***
### Question 1

# This script is importing and tidying the Johnson Elementary Project Data.

```{r}
library(tidyverse)
library(here)
```

```{r}
johnson.data <- read_csv("Johnson Elementary ELA Project.csv",  skip = 4,
                              col_names = c("Room", "ID", "Tardies;Absences",
                                            "Q1", "Q2", "Q3", "Q4"))

```

```{r}
tidyjohnson.data <- johnson.data%>%
  fill(Room)%>%
  drop_na(ID) %>%
  separate('Tardies;Absences', into = c("tardies", "absences"), sep = ";", 
           convert = TRUE)%>%
  gather(`Q1`, `Q2`, `Q3`, `Q4`,
         key = "Quarters",
         value = "Ratings")
```

***
### Question 2

# This script is providing summary statistics for the number of tardies and absences for each room.

```{r}
tard.abs.summaries <- tidyjohnson.data %>%
  group_by(Room) %>%
  summarize(tardies_mean = mean(tardies, na.rm = TRUE),
            tardies_sd = sd(tardies),
            absences_mean = mean(absences, na.rm = TRUE),
            absences_sd = sd(absences))
```

***
### Question 3

# This script is to show the frequency of missing data points for each quarter by room

```{r}
missing.room.data <- tidyjohnson.data %>%
  group_by(Room, Quarters) %>%
  summarize(sum(is.na(Ratings)))

missing.room.data
```

***
### Question 4

# This is a graph that illustrates the median rating trends across the four quarters forthe fifth-grade classes

```{r}
median.chart <- tidyjohnson.data %>%
  group_by(Room, Quarters) %>%
  summarize(median_scores = median(Ratings, na.rm = TRUE)) %>%
  ggplot(aes(x = Quarters, y = median_scores, group = Room)) +
  geom_point(aes(color = Room)) +
  facet_wrap(vars(Room)) +
  guides(color = "none")
    labs(x = "Quarters", y = "Median Ratings", 
         title = "Fifth Grade Median Ratings across Four Quarters for each Classroom")

median.chart
```

***
### Question 5

# This is a graph that compares Airline carriers in terms of the proportion of 2013 flights that arrived on time when traveling from New York City to Atlanta.

```{r}
flights <- read_csv(here("Data","NYC Flights 2013.csv"))
```

```{r}
flights %>% mutate(on_time = (arr_delay <= 0)) %>%
  filter(dest == "ATL", !is.na(on_time)) %>%
  ggplot(aes(x = carrier)) +
    geom_bar(aes(fill = on_time), position = "fill") +
    labs(title = "On-Time Rates for Carriers from NYC to ATL",
         x = "Airline Carrier",
         y = "Proportion") +
  coord_flip() +
    scale_fill_discrete(name = "Arrival Status",
                      labels = c("Not on Time",
                                 "On Time")) +
    scale_x_discrete(labels = c("United", "Delta",
                              "ExpressJet", "Envoy", 
                              "Southwest", "AirTran", 
                              "Endeavor")) 
```

